---
title: "Project - Part2"
author: "Anusha Dasari"
output:
  pdf_document: default
  html_document:
    df_print: paged
---
*Loading the required libraries:*

```{r}
library(readxl)
```

*Loading LINTHALL data for analysis:*
```{r data}
data_2=read_excel("/Users/anusha_dasari/Downloads/LINTHALL.xlsx")
data_2=data_2[-1:-3]
data_2
```
```{r cor}
# Correlation Matrix of LINTHALL data
df=data_2[-1]
df_cor=cor(df)
```
```{r eigenvector}
# Eigen vectors
ev<-eigen(df_cor)
V=ev$vectors
K=ev$values
```


```{r stand}
# Standardizing the data
data_scale<-scale(data_2)
data_scale = as.data.frame(data_scale)
```

*Performing linear regression on standardized data:*
```{r std}
std_model <- lm(BIO~H2S+SAL+Eh7+pH+BUF+P+K+Ca+Mg+Na+Mn+Zn+Cu+NH4, data=data_scale)
coef(std_model)
```

Above coefficients are estimates of theta . Regression coefficients of original data can be directly calculated using the formula:
$$\hat{\beta_{j}}= \frac{S_y}{S_j}*\hat{\theta_{j}}$$
Sy = Standard Deviation of  Y.
Sj = Standard Deviation of  Xj

(This will be calculated later in this report.) 

```{r pc}
# Finding principle components
PC= as.data.frame(as.matrix(data_scale[-1])%*%V)
PC$Y=data_scale$BIO
names(PC)=c('C1','C2','C3','C4','C5','C6','C7','C8','C9','C10','C11','C12','C13','C14','Y_tilde')
PC
```

```{r pcr}
# Performing principle component regression
pcr_full=lm (Y_tilde~C1+C2+C3+C4+C5+C6+C7+C8+C9+C10+C11+C12+C13+C14,data=PC)
alpha=coef(pcr_full)
alpha
```

```{r theta}
theta= V %*% alpha[-1]
theta
```



```{r components}
s = 'Y_tilde~C1'
p = length(theta)
theta_mat = matrix(nrow=p, ncol=(2+p))
i=1
while(i<=p){
  pcr = lm(as.formula(paste(s,'-1')),data=PC)
  alpha = as.matrix(coef(pcr))
  theta= V[,1:i]%*%alpha
  r2 = summary(pcr)$r.squared
  theta_mat[i,1] = i
  theta_mat[i,2] = r2
  theta_mat[i,3:(p+2)]= theta
  s = paste(s,sprintf('+C%d',i+1))
  i=i+1}
theta_mat = as.data.frame(theta_mat)
names(theta_mat) = c('ncomp', 'R^2', paste('theta',1:p,sep=''))
theta_mat
```
We use principal components C1,C2,C3,C4,C5,C6,C7,C8,C9,C10 and C11 in the model as per the Rsquare results in above table. 

Now lets calculate the regression coefficients of the original data using theta values generated previously.
Formula used:

$$\hat{\beta_{j}}= \frac{S_y}{S_j}*\hat{\theta_{j}}$$
Regression Coefficients of the original model:
```{r beta}
beta1=(sd(data_2$BIO)/sd(data_2$H2S))*theta[1]
beta2=(sd(data_2$BIO)/sd(data_2$SAL))*theta[2]
beta3=(sd(data_2$BIO)/sd(data_2$Eh7))*theta[3]
beta4=(sd(data_2$BIO)/sd(data_2$pH))*theta[4]
beta5=(sd(data_2$BIO)/sd(data_2$BUF))*theta[5]
beta6=(sd(data_2$BIO)/sd(data_2$P))*theta[6]
beta7=(sd(data_2$BIO)/sd(data_2$K))*theta[7]
beta8=(sd(data_2$BIO)/sd(data_2$Ca))*theta[8]
beta9=(sd(data_2$BIO)/sd(data_2$Mg))*theta[9]
beta10=(sd(data_2$BIO)/sd(data_2$Na))*theta[10]
beta11=(sd(data_2$BIO)/sd(data_2$Mn))*theta[11]
beta12=(sd(data_2$BIO)/sd(data_2$Zn))*theta[12]
beta13=(sd(data_2$BIO)/sd(data_2$Cu))*theta[13]
beta14=(sd(data_2$BIO)/sd(data_2$NH4))*theta[14]

betas=c(beta1,beta2,beta3,beta4,beta5,beta6,beta7,beta8,beta9,beta10,beta11,beta12,beta13,beta14)

betas
```

Intercept of the original model:
```{r intercept}
bi<-c()
for(x in 1:14){
  bi[x]=betas[x]*(sum(data_2[x+1])/44)
}
interceptEsti = mean(data_2$BIO) - sum(bi)
interceptEsti
```
---------------------------------------------