---
title: "Project - Part1"
author: "Anusha Dasari"
date: "11/26/2021"
output:
  pdf_document: default
  html_document:
    df_print: paged
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


*Loading the required libraries:*

```{r}
library(readxl)
library(olsrr)
library(GGally)
library(ggplot2)
```
*Loading LINTHALL data for analysis:*
```{r data}
data=read_excel("/Users/anusha_dasari/Downloads/LINTHALL.xlsx")
data=data[-1:-3]
data
```

*Performing Ordinary Least Square Regression and shown below are the estimated regression coefficients:*
```{r ols}
model_ols<-lm(BIO~H2S+SAL+Eh7+pH+BUF+P+K+Ca+Mg+Na+Mn+Zn+Cu+NH4, data=data)
coef(model_ols)
```

*Performing collinearity check:*
```{r scatter, fig.height=19,fig.width=19}
ggpairs(data[-1])
```


```{r cor}
corMat=cor(data[-1])
corMat
```

From the above scatter plot and correlation matrix result, we can say that there is collinearity between following predictors:

1. BUF and pH,NH4 
2. pH and Ca
3. Mg and  K, Na

We can confirm this by further analysis:

```{r eigenvalues}
ev<-eigen(corMat)
eigenvalues=ev$values
sum(1/eigenvalues)>(5*14)
```
There is a condition that proves that collinearity exists in the data, i.e. if the sum of reciprocals of eigen values is greater than five times the number of predictor variables there exists collinearity. Since it is true in this case, collinearity exists! Lets do further diagnostics.


*Performing collinearity diagnostics to know more about collinearity:*
```{r collinearity}
collDiag=ols_coll_diag(model_ols)
collDiag
```

```{r CI}
#sets of collinearity
sum(collDiag$eig_cindex$`Condition Index`>15)
```
If the condition indices is small (<15), then predictor variables are not collinear. But from this results we can say that there are 9 sets of predictors shows collinearity as their condition indices>15. 
```{r pred}
#predictors effected by collinearity:
vifvals=collDiag$vif_t$VIF
which(vifvals>10)
```
From the above results, we can see that for variables pH, BUF, Ca, Mg, Na, Zn VIF> 10. This indicates that these predictors are effected by collinearity.


Therefore we proved that collinearity exists and as per the analysis the variables that contribute to collinearity are pH, BUF, Ca, Mg, Na, Zn.

-----------------------------------


